---
title: "Lab 02 - Plastic waste"
author: "AllisonLi"
date: "01252025"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
library(datasauRus) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1.1

```{r plastic-waste-continent}
# the distribution of plastic waste per capita in 2010
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, color = continent, fill = continent)) +
  geom_histogram(binwidth = 0.2)+
  facet_wrap(~ continent)
 
# identify the outlier country: 
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

In general, three out of the six continents show less than .5 plastic waste per capita. Asia and North America, as well as an outlier in South America show more relatively more plastic waste per capita. Additionally, there is an outlier in North American, indicating that Trinidad and Tobago has more than 3.5 plastic waste per capita. The internet suggests that is probably due to single-use plastic bottles and small population size. 

### Exercise 2.1
```{r plastic-waste-density}
# density plot
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.5)
```

### Exercise 2.2
Alpha is known as transparency, it makes the variable either less or more transparable. Color and fill are usually a way to assign a color or a certain aesthetic to a variable. For alpha, it is a setting characteristic that is assigned as a constant value for the whole layer. All the variables presented will be set to the same alpha value. In other words, Alpha does not present the differences in data. 

```{r plastic-waste-boxplot}
# side-by-side box plots
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
  geom_boxplot()
```

### Exercise 3.1

```{r plastic-waste-violin}
# Violin plots
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_violin()
```

Box plot clearly shows the median value for each continent and each outlier, while the violin plot shows the density, indicating how many countries there are for each level of plastic waste per capita in each continent. For example, the violin plot allows us to see most countries in Africa has close to 0 plastic waste per capita. The median value and outlier are apparent in the box plot. 

### Exercise 4.1

```{r plastic-waste-scatterplots}
# scatter plots
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = mismanaged_plastic_waste_per_cap,
    y = plastic_waste_per_cap
  )
) +
  geom_point()
```

It seems to be that most dots (countries) appeared to be around the low value. Although the relationship does not look like a linear relationship, it could be considered as a slightly positive linear relationship. This suggested that as the amount of mismanaged plastic waste per capita increases, the plastic waste per capita also increases. 

### Exercise 4.2

```{r plastic-waste-mismanaged-continent}
# colored scatter plots
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = mismanaged_plastic_waste_per_cap,
    y = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_point()
```

Based on the graph, it seems like there isn't a clear distinction for each continente in terms of the relationship between the plastic waste per capita and mismanaged plastic waste per capita. 

### Exercise 4.2

```{r plastic-waste-population-total}
# plastic waste per capita and total population
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    y = total_pop,
    color = continent,
    fill = continent
  )
) +
  geom_point()

# plastic waste per capita and coastal population
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    y = coastal_pop,
    color = continent,
    fill = continent
  )
) +
  geom_point()
```

None of both seems to have a strong linear relationship. However, the relationship between plastic waste per capita and coastal population seems to be more strongly linearly associated. 

### Exercise 5

```{r recreate-viz}
# graph
plastic_waste %>%
  filter(plastic_waste_per_cap <= 3.5) %>%
  mutate(costal_pop_pro = coastal_pop / total_pop) %>%
    ggplot(
    mapping = aes(
      x = costal_pop_pro,
      y = plastic_waste_per_cap,
      color = continent,
    )
  ) +
    geom_point() +
    geom_smooth(color = "black", method = loess, se = TRUE)
    labs(caption = "Coastal population proportion")

```

5.1. Recreate the following plot, and interpret what you see in context of the data.

This graph shows a general trend of the data that indicates there is no linear relationship between the two variables and there confidence interval is large. In general, it seems that countries with higher coastal population proportion have higher plastic waste production per capita.
